---
title: "Class 8"
author: "Brian Valadez"
date: "October 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Our first example with **kmeans()** function

```{r}
#example plot to see how Rmarkdown works
plot(1:10, typ="l")

```
#back to kmeans
```{r}

```
#our own work
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)

```
```{r}
k <- kmeans(x, centers=2, nstart=20)
k
```


```{r}
k$cluster
```
```{r}
table(k$cluster)
```

```{r}
k$centers
```
```{r}
palette(c("blue","red"))
plot (x, col=k$cluster)
points(k$centers, col="green", pch=20, cex=3)
```

```{r}
k3 <- kmeans(x, centers = 3, nstart = 20)
k3$tot.withinss
k$tot.withinss
#the higher the k the lower  our value
```

```{r}
k2 <- kmeans(x, centers = 2, nstart = 20)
k3 <- kmeans(x, centers = 3, nstart = 20)
k4 <- kmeans(x, centers = 4, nstart = 20)
k5 <- kmeans(x, centers = 5, nstart = 20)
k2$tot.withinss
k3$tot.withinss
k4$tot.withinss
k5$tot.withinss
plot(c(k2$tot.withinss,
k3$tot.withinss,
k4$tot.withinss,
k5$tot.withinss))
```


 Use the kmeans() function setting k to 2 and nstart=20
 Inspect/print the results
 Q. How many points are in each cluster?
 Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?
 Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 Q. Repeat for k=3, which has the lower tot.withinss? 
 
 
 
 
 
 
 
 

 #Hierarchical clustering in R
Trying out the **hclust()** fuction for hierarchical clustering in R 
```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

```{r}
plot(hc)
#draw a line on the dendrogram
abline(h=8, col="red")
#cut the tree to yield cluster membership
cutree(hc, h=8)
```
 
 
 
 ### A more 'real' example of data
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("red","blue","green"), each=50) )
plot(x, col=col)

plot(hclust(dist(x)))

```
use cutree()function to return 2 and 3 clusters
```{r}
group.3 <- cutree(hc, k=3)
group.2 <- cutree(hc, k=2)
group.2
group.3

```
compare to known groups
```{r}
table(group.3, col)
```


# Principal comment analysis
```{r}
mydata <- matrix(nrow=100, ncol=10) 
rownames(mydata) <- paste("gene", 1:100, sep="") 
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
 paste("ko", 1:5, sep="") ) 
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1)) 
 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
pca <- prcomp(t(mydata), scale=TRUE)
```

not the prcomp()function wants us to take the transpose of our data so we fliped rows and collums
```{r}
head(t(mydata))
```
now to find the "structure in this data with **prcomp()**
 
 
```{r}
pca <- prcomp(t(mydata), scale=TRUE)
pca

```
making the pc plot
```{r}
plot(pca$x[,1], pca$x[,2]) 
#square root of variance gives standard deviation
pca.var <- pca$sdev^2 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per

```
making a screen plot to find variation in the data
```{r}

barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```
making the graph more useful
```{r}
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
#now we are addind labels
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata)) 
```

Hands on section 
```{r}
library(readr)
UK_foods <- read_csv("UK_foods.csv")
dim(UK_foods)
head(UK_foods)

knitr::kable(UK_foods, caption="The full UK foods data table")
pca <- prcomp( t(x) )
summary(pca)
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(UK_foods))
```

 
 
 
 
 
 
 