---
title: "Bimm 143 Class 13"
author: "Brian Valadez"
date: "November 13, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Get HIV-Pr structure from PDB database
 We are working with 1HSG
```{r}
library(bio3d)
HSG.pdb <- get.pdb("1hsg")
```
Read this in to R and examine composition 
```{r}
hiv <- read.pdb("1hsg")
hiv
```
Split in to protein and ligand
```{r}
protein <-trim.pdb(hiv, "protein") 
ligand <- trim.pdb(hiv, "ligand")
```
```{r}
protein
```
```{r}
write.pdb(protein, file="1hsg_protein.pdb")
```
```{r}
ligand
```

```{r}
write.pdb(ligand,"1hsg_lligand.pdb")
```
## Docking with vina
We run this comand 
"\Program Files (x86)\The Scripps Research Institute\Vina\vina.exe" --config config.
txt --log log
## Read docking results
we got a file all.pdbqt that we need to making into a PDB format
```{r}
res <- read.pdb("all.pdbqt", multi=T)
write.pdb(res, "results.pdb")
```
```{r}
res

```
Checking to see which is best ( should be the first one)
```{r}
ori <- read.pdb("1hsg_lligand.pdbqt")
rmsd(ori,res)
```
## Section 3.1
```{r}
HELpdb <- read.pdb("1HEL")
modes <- nma(HELpdb)
plot(modes, sse=HELpdb)
```



```{r}
mktrj(modes, mode=7, file="nma_7.pdb")
```

